The impacts of park access on health outcomes: a spatial comprehensive approach

### 2019-02-11 manuscript.Rnw
Lock down coefficients on size and distance.

Numerous previous studies have found relationships between aspects of the built
environment and aspects of physical, mental, social, and economic health. Many
of these findings are summarized in Table \ref{tab:parkbenefits}.
For the purposes of this study, it is clear that parks and other
vegetated areas support outdoor physical activity in ways distinct from other
urban environments \citep{Giles-Corti2005, Wells2007}.

@@

While several of the aforementioned studies yielded results consistent with the
hypothesis that park access and use confer health benefits stemming from
increased physical activity and attendant decreases in obesity, all of the
studies -- except some included in the meta-analysis by \citet{Bancroft2015} -–
were conducted at the level of the city or metropolitan statistical area.
Metropolitan-level analyses do not capture the within-city variation in
accessibility to parks that exists. Some cities with large amounts of total
greenspace may nevertheless inequally distribute this space throughout the city
leading to areas with poor access. Conversely, a city with a smaller overall
proportion of greenspace may give all of its citizens better access to significant
greenspace. Considering access at the neighborhood level within a single city
may eliminate some regional or cultural fixed effects affecting
metropolitan-level analyses.

@@

A key theoretical outcome of random utility choice models is that the
consumer surplus, or the total value of all alternatives of the choice set, can
be obtained as the log-sum of the denominator of the choice probability equation
\citep{Small1981}:
\begin{equation}\label{eq:cs}
CS_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
\end{equation}
There are several advantages to a log-sum defined metric relative to
buffer-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary cutoff asserted by the researcher. This allows for the fact that some
people are more or less sensitive to distances, and that distance is a
continuous, and not a binary, phenomenon. Second, the random utility formulation
allows the researcher to include any attribute of the park; in this case, we
consider the size of the park as an element of accessibility. This suggests that
not all parks are equal, and that a large park such as Central Park on Manhattan
may provide health and activity benefits over a much larger area than a smaller
community square.

@@

In spite of its advantages, logsum-based accessibilities have not received
as much application in the accessibility literature as distance-based or even
gravity-based measures. They are commonly used, however, in alternatives analyses of
transit infrastructure improvements \citep{DeJong2007}. A reason for this is
likely that such analyses are usually conducted in the context of a travel
demand model, where calibrated and multimodal logsums are readily available
\citep{Geurs2010}.

\section{Empirical Methodology}
In this section we describe an experiment with data for New York City, where
we compute a holistic logsum-based accessibility component and model the
relationship between this measure and physical activity rates, controlling for
spatial effects and socioeconomic factors.

@@

The primary dataset is a geographic polygons shapefile of Census tracts in
New York City. We appended relevant sociodemographic data for each tract level
from the American Community Survey 2011-2015 5-year estimates. Data on physical
activity participation rate and obesity rates for each tract are available from
the Centers for Disease Control and Prevention’s 500 Cities Project.
After removing tracts with missing population information or outlying and unusual
characteristics, we have Sexpr{nrow(tracts)} complete cases. Table
\ref{tab:tractsdata} presents key descriptive statistics for these data.

@@

In a destination choice framework, the tracts represent the "origins" and
"destinations" are parks in New York City.
We retrieved a shapefile of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance. Upon inspection,
we removed several facilities that do not qualify as publicly accessible green
space, such as Yankee Stadium and its surrounding parking lots. We also removed
parks of less than half an acre in size, as these appear to be predominantly
vacant lots or parking strips rather than legitimate public green space. This 
leaves us with \Sexpr{nrow(parks)} separate parks.

@@

bind_rows(tract_summary, park_summary) %>%
  kable(escape = FALSE, digits = 2, 
        table.envir = "table*", booktabs = TRUE,
        caption = "Descriptive Statistics of Tract and Park Variables",
        col.names = c(" ", "Minimum", "Median (IQR)", "Maximum")) %>%
  kable_styling() %>%
  group_rows(paste("Tract Variables, N =", nrow(tracts@data)), 1, 13) %>%
  group_rows(paste("Park Variables, N =", nrow(parks)), 14, 15)

@@

We predict the rate of physical activity ($y$) in a census tract as a
function of the tract's sociodemographic characteristics and choice-based
accessibility to parks with a semi-log specification:
\begin{equation}\label{eq:themodel}
\log(y_i) = f(\vect{x}_i, CS_i, g(W, \mathit{Obesity}, \vect{X}))
\end{equation}
<<base_formula>>=
base_formula <- formula(
  ~ log(density) +
    lowincome + 
    fulltime + college + single +
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
    black + asian + hispanic + other)

termnames <- c("log(density)" , "log(Income)", 
               "Full-time workers", "College-educated",  "Single Adults", 
               "Share of children (0-17)", "Share of young adults (18-29)", 
               "Share of seniors (65+)", "Black population share", "Asian population share", 
               "Hispanic population share", "Other Minorities")
@
where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes,
$CS_i$ is a choice-based
accessibility measure for the tract computed from Equation \ref{eq:cs} and
$g(W, \vect{y}, X)$
represents the impact of spatial neighborhood effects on the outcome.

Using tracts as the unit of observation in the model allows us to consider the
availability of parks at a sub-metropolitan level, but it also introduces the problem
of spatial correlation of unobserved errors and spatial dependence in the data
generating process \citep{Cliff1970, Anselin1980}. Spatial correlation occurs
when spatially-distributed unobservable or missing variables (school quality,
neighborhood prestige, etc.) influence the modeled outcomes. Spatial dependence
occurs when the outcome {\em depends} on the outcome in neighboring areas; for
example, seeing neighbors exercising may encourage exercise. The relaxed spatial
Durbin model (SDM) proposed by \citet{Burridge1981} controls for both processes:
\begin{equation}\label{eq:sdm}
  \vect{y}= \rho W\vect{y} + X\vect{\beta} + WX\vect{\gamma} + \vect{\epsilon}
\end{equation}
where $\rho$ is the estimated spatial correlation between the outcome variables
$y_i, y_j, \ldots  \in \vect{y}$, $W$ is a matrix specifying the spatial
relationship between observations $i$ and $j$ in the dataset,
$\vect{\beta}$ is a vector of estimable coefficients relating the attributes $X$
of an observation to its outcome, and $\vect{\gamma}$ is a vector of
estimated coefficients relating the attributes of an observation's {\em neighbors}
to its outcome.

@@

It may  be more efficient to estimate a restricted spatial autoregressive
model -- that controls for only correlation or only dependence -- and therefore
save degrees of freedom in the estimation.  \citet{tra_autoregression_paper}
illustrate a decision algorithm originally proposed by \citet{Florax2003} which uses
a likelihood ratio test of the SDM against a restricted spatial error model (SEM):
if the likelihood ratio test rejects that the models are equivalent, then the SDM
should be used.

A final point to note about spatial autoregressive models is that the partial
derivative of the model with respect to a single regressor $\vect{x_k} \in X$
does not equal the estimated coefficient $\beta_k$. Consequently, the estimated
coefficients do not represent the {\em effect} of a variable on the outcome. The
direct effects of an observation's attribute on its outcome, the indirect effects
of neighbors' attributes on an outcome, and the total effect of own and neighbors'
attributes can attained via Monte Carlo simulation. In this study we use the
\texttt{spdep} package for R \citep{Bivand2013} for model estimation and impact
simulation.

@@

We initially estimated potential values of the $\beta$ coefficients by
applying a limited-memory, box-constrained optimization algorithm maximizing the
SDM model log-likehood. The constraint required that the coefficient on
size be positive and that on distance negative; all else equal, we assert people
will prefer to use larger parks and parks that are nearer to their residence. This
resulted in $\beta_d = -0.42$  and $\beta_s = 0.000$. Given our expectation that
park size is a valuable attribute, we subsequently adjusted this value to
a value of $1.00$. This implies that each mile

@@

<<logsum_betas>>=
tracts@data <- tracts@data %>%
  mutate(
    access_ls = calculate_park_logsums(log(distances), parks$log_size,
                                       betas = c(-1.79, 0)),
    tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
                                       betas = c(-0.79, 0, 0.1)),
    access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
    tweets_ls = (tweets_ls - mean(tweets_ls) )/ sd(tweets_ls)
  )

@@

<<map, echo=FALSE, message=FALSE, warning=FALSE>>=
library(ggmap)
map <- ggmap(get_map(unname(st_bbox(tracts_sf)), source = "stamen", maptype = "toner-lite"),
             extent = "device")
map +
  geom_sf(data = tracts_sf, inherit.aes=FALSE,
          aes(fill = cut(access_ls, breaks = c(-Inf, -3, -2, -1, 0, 2, 3, Inf))),
          alpha = 0.8) +
  scale_fill_brewer("Relative Park Logsum", palette = "GnBu")
@
  \caption{Relative surplus-based accessibility in New York City}
  \label{fig:map}
\end{figure*}

\section{Results}

<<spatial selection>>=
physact_base_sem <- errorsarlm(update(base_formula, physact ~ .),
                             data = tracts@data, listw = W, zero.policy = TRUE)
physact_base_sdm <- lagsarlm(update(base_formula, physact ~ .),
                           data = tracts@data, listw = W, zero.policy = TRUE,
                           type = "mixed")
test_sdmsem <- lmtest::lrtest(physact_base_sem, physact_base_sdm)
@

<<access_sdm>>=
# estimate models
pa_access_sdm <- update(physact_base_sdm, .~ . + access_ls)
pa_tweets_sdm <- update(physact_base_sdm, .~ . + tweets_ls)
@

We estimated SDM and SEM models regressing the physical activity rate against
against the base covariates (without the accessibility logsum). A likelihood
ratio test rejects that the SEM is equivalent
($\chi^2 = \Sexpr{test_sdmsem$Chisq[2]}$ on \Sexpr{test_sdmsem$Df[2]} degrees of
freedom), so we use the SDM specification only going forward. Table
\ref{tab:pa_models} presents the estimated coefficients for the base model as
well as models including logsum-based accessibility with and without Twitter
activity. A few initial observations can be made from this table. First, adding
the logsum measurement significantly improves the overall model fit, using a
likelihood ratio test on the model log-likelihood statistics. Second, the
coefficient estimates on the controlling variables remain virtually unchanged
from the base model.

\begin{table*}
  \caption{Estimated Coefficients Relating Park Access to Physical Activity}
  \label{tab:pa_models}
<<pamodelchart, results="asis">>=
pa_models <- list(Base = physact_base_sdm, `Access` = pa_access_sdm,
                       `Access + Tweets` = pa_tweets_sdm)
texreg(pa_models, digits = 4, table = FALSE, single.row = TRUE,
       booktabs = TRUE, use.packages = FALSE, 
       custom.coef.names = c("Intercept", termnames, "$\\lambda$",
                             "Access logsum", 
                             "Access + Tweets logsum"))
#' screenreg(pa_models, digits = 4, table = FALSE)
@
\end{table*}

Recall from the presentation of spatial econometrics above that the total effect of a
covariate on the dependent variable is different from the estimated coefficient
and involves interaction between direct and indirect effects, or between
the $\beta$ and $\gamma$ parameters. This is particularly important when the
$\beta$ and $\gamma$ estimates for a particular covariate show a different
sign, such as in the proportion of black residents in a census tract, as
interpreting the direction and magnitude of the relationship is not straightforward.

<<extract_impacts>>=
#' Get the MC simulated impact coefficients and significance from a model
#' @param sdm An autoregressive lag (SAR or SDM) model object `lagsarlm`
#' @return A data frame with the effect (direct, indirect, total), variable,
#'   simulated mean impact, and significance p-value
extract_impacts <- function(sdm) {

  impacts_summary <- summary(
    impacts(sdm, tr=trMC,  R = 1000, useHESS = !is.logical(sdm$fdHess)),
    zstats = TRUE)

  estimate <- list(
    Direct = impacts_summary$direct_sum,
    Indirect = impacts_summary$indirect_sum,
    Total = impacts_summary$total_sum
  ) %>%
  lapply(function(s) {
    data_frame(
      term = names(s$statistics[,1]),
      estimate = s$statistics[,1],
      std.error = s$statistics[,2]
    )
  }) %>%
    bind_rows(.id = "effect")

  pval <- as_data_frame(impacts_summary$pzmat) %>%
    mutate(
     term = rownames(impacts_summary$pzmat)
    ) %>%
    gather("effect", "p-val", -term)

  left_join(estimate, pval, by = c("effect", "term"))
}
@

\begin{table*}
  \caption{Simulated Impact of Accessibility on Physical Activity}
  \label{tab:pa_impacts}
<<impacts, results="asis">>=
physact_impacts <- pa_models %>%
  lapply(extract_impacts) %>%
  bind_rows(.id = "model") %>% ungroup() %>%
  rename(submodel = effect)

effect_estimate <- physact_impacts %>%
  filter(model == "Access + Tweets", term == "tweets_ls", submodel == "Total") %>%
  pull(estimate)

density_estimate <- physact_impacts %>%
  filter(model == "Access + Tweets", term == "log(density)", submodel == "Total") %>%
  pull(estimate)

physact_impacts %>%
  transmute(
    model, term, submodel, estimate,
    output = paste(round(`p-val`, 5),
                   gtools::stars.pval(`p-val`))
  ) %>%
  filter(term %in% c("access_ls", "tweets_ls"), submodel == "Total") %>%
  dplyr::select(-term, -submodel) %>%
  kable("latex", booktabs = T, escape = FALSE,
        col.names = c("Logsum Measure", "Estimate", "$p$-value")) %>%
  footnote("$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$", escape = F)
@
\end{table*}

Table \ref{tab:pa_impacts} shows the simulated marginal impact of the
accessibility logsum measurements on the physical activity rate in each tract.
Given that the logsum is a relatively continuous measure over space, and that it
would be difficult to improve the park accessibility of a tract without also
improving the accessibility of the tract's neighbors, the total effect combining
the direct effect and the indirect effects of all neighbors is
most appropriate. The total effect of tract-level accessiblity on physical
activity rates is strongly significant, regardless of whether we consider Twitter
activity.

It is of course necessary to also consider the magnitude of the effect.
As this is a semi-log model, the coefficient implies a unit increase
in the access and tweets logsum leads to a
\Sexpr{round(effect_estimate * 100, 3)}\% increase in the physical activity rate.
Referring to the map of logsum measures in Figure \ref{fig:map}, the measures
span from roughly $-3$ in heavily industrialized segments of southern Brooklyn
to roughly $3$ in tracts immediately adjacent to Central Park on the Upper East
Side. Thus moving from the least accessible parts of the city to the most
accessible will reduce the expected physical activity rate by
\Sexpr{round(effect_estimate * 100 * 6, 3)}\%. In other words, if a tract with 
poor accessibility is expected to achieve the sample mean of 
\Sexpr{round(mean(tracts$physact), 1)}\%, the same tract with excellent accessibility
is expected to achieve a rate of 
\Sexpr{round(mean(tracts$physact) * (1 + effect_estimate  * 6), 2)}\%. 
For an aggregate statistic strongly influenced by other socioeconomic factors, 
this is not an insubstantial marginal effect.

<<obesity_models>>=
obese_access_sdm <- update(pa_access_sdm, obesity ~ .)
obese_physact_sdm <- update(obese_access_sdm, . ~ . + physact)
@

We now consider the impacts of a model where the dependent variable is 
the {\em obesity} rate, and physical activity becomes an independent covariate
alongside the controlling variables and the tweets-augmented logsum variable. 
Table \ref{tab:obese_impacts} presents the simulated impacts resulting from
this model.

\begin{table*}
  \caption{Simulated Impacts of Variables on Obesity}
  \label{tab:obese_impacts}
<<tab:obese_impacts, results="asis">>=
extract_impacts(obese_physact_sdm) %>%
  group_by(effect) %>%
  transmute(
    term,
    output = paste(round(`estimate`, 5),
                   "(",round(`std.error`, 5),")",
                   gtools::stars.pval(`p-val`)),
    order = row_number()
  ) %>%
  spread(effect, output) %>%
  arrange(order) %>%
  dplyr::select(-order) %>%
  mutate(
    term = c(termnames, "Access + Tweets logsum", "log(Physical Activity Rate)")
  ) %>%
  kable("latex", booktabs = T, escape = FALSE, digits = 3, 
        col.names = c("", "Direct", "Indirect", "Total")) %>%
  footnote("$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$", escape = F)
@
\end{table*}

>>>>

Numerous previous studies have found relationships between characteristics of the built
environment and aspects of physical, mental, social, and economic health. Many
of these findings are summarized in Table \ref{tab:parkbenefits}.
For the purposes of this study, it is clear that parks and other
vegetated areas support outdoor physical activity in ways distinct from other
urban environments \citep{Giles-Corti2005, Wells2007}.

@@

While several of the aforementioned studies reported results consistent with the
hypothesis that park access and use confer health benefits stemming from
increased physical activity and attendant decreases in obesity, all of the
studies -- except some included in the meta-analysis by \citet{Bancroft2015} -–
were conducted at the level of the city or metropolitan statistical area.
Metropolitan-level analyses do not capture the within-city variation in
accessibility to parks that exist in many cities. Some cities with large amounts of total
greenspace may nevertheless inequally distribute this space throughout the city,
leading to areas with poor access. Conversely, a city with a smaller overall
proportion of greenspace may give all of its citizens better access to sufficient
greenspace to meet their needs or wants for physical activity. Considering access at the neighborhood level within a single city
may also eliminate some regional or cultural fixed effects affecting
metropolitan-level analyses.

@@

A key theoretical outcome of random utility choice models is that the
consumer surplus, or the total weighted value of all alternatives of the choice set, can
be obtained as the log-sum of the denominator of the choice probability equation
\citep{Small1981}:
\begin{equation}\label{eq:cs}
CS_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
\end{equation}
There are several advantages to a log-sum defined metric relative to
buffer-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary cutoff asserted by the researcher. This allows for the fact that some
people are more or less sensitive to distances, and that distance is a
continuous, non-binary phenomenon. Second, the random utility formulation
allows the researcher to include any attribute of the park; in this case, we
consider the size of the park as an element of accessibility. This suggests that
not all parks are equal, and that a large park such as Central Park on Manhattan
may provide health and activity benefits over a much larger area than a smaller
community square.

@@

In spite of its advantages, logsum-based accessibilities have not received
as much application in the accessibility literature as distance-based or even
gravity-based measures. They are commonly used, however, in alternatives analyses of
transit infrastructure improvements \citep{DeJong2007}. A reason for this is
likely that such alternatives analyses are usually conducted in the context of a travel
demand model, where calibrated and multimodal logsums are readily available
\citep{Geurs2010}.

\section{Empirical Application}
In this section we describe an experiment with data for New York City, where
we compute a holistic logsum-based accessibility to parks metric and model the
relationship between this measure and physical activity rates, controlling for
spatial effects and socioeconomic factors. We subsequently model the effect
of this accessibility on obesity.

@@

The primary dataset is a geographic polygons shapefile of Census tracts in
New York City. We appended relevant sociodemographic data for each tract level
from the American Community Survey 2011-2015 5-year estimates. Data on physical
activity participation rate and obesity rates for each tract are available from
the Centers for Disease Control and Prevention’s 500 Cities Project (WE NEED A CITATION).
After removing tracts with missing population information or outlying and unusual
characteristics, we have Sexpr{nrow(tracts)} complete cases. Table
\ref{tab:tractsdata} presents key descriptive statistics for these data.

@@

In a destination choice framework, the tracts represent the "origins" and
"destinations" are parks in New York City.
We retrieved a shapefile of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance (WE NEED A CITATION).
Upon inspection,
we removed several facilities that do not qualify as publicly accessible green
space, such as Yankee Stadium and its surrounding parking lots. We also removed
parks of less than half an acre in size, as these appear to be predominantly
vacant lots or parking strips rather than legitimate public green space. This
leaves us with \Sexpr{nrow(parks)} parks. It is worth noting that the particular
dataset classifies specific facilities within larger parks as independent
features: baseball fields within Central Park, for example, are themselves
unique greenspace. We elect to retain these independent features as they may
generate additional activity.

@@

bind_rows(tract_summary, park_summary) %>%
  kable(escape = FALSE, digits = 2,
        table.envir = "table*", booktabs = TRUE,
        caption = "Descriptive Statistics of Tract and Park Variables",
        col.names = c(" ", "Minimum", "Median (IQR)", "Maximum")) %>%
  kable_styling() %>%
  group_rows(paste("Tract Variables, N =", nrow(tracts@data)), 1, 16) %>%
  group_rows(paste("Park Variables, N =", nrow(parks)), 17, 18)

@@

We predict the rate of physical activity ($y$) in a census tract as a
function of the tract's sociodemographic characteristics and choice-based
accessibility to parks with the following  specification:
\begin{equation}\label{eq:themodel}
 y_i = f(\vect{x}_i, CS_i, g(W, \vect{y} \vect{X}))
\end{equation}
where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes,
$CS_i$ is a choice-based
accessibility measure for the tract computed from Equation \ref{eq:cs} and
$g(W, \vect{y}, X)$
represents the impact of spatial neighborhood effects on the outcome.

Using tracts as the unit of observation in the model allows us to consider the
availability of parks at a sub-metropolitan level, but it also introduces the problem
of spatial correlation of unobserved errors and spatial dependence in the data
generating process \citep{Cliff1970, Anselin1980}. Spatial correlation occurs
when spatially-distributed unobservable or missing variables (school quality,
neighborhood prestige, etc.) influence the modeled outcomes. Spatial dependence
occurs when the outcome {\em depends} on the outcome in neighboring areas; for
example, seeing neighbors exercising may encourage exercise. The relaxed spatial
Durbin model (SDM) proposed by \citet{Burridge1981} controls for both processes:
\begin{equation}\label{eq:sdm}
  \vect{y}= \rho W\vect{y} + X\vect{\beta} + WX\vect{\gamma} + \vect{\epsilon}
\end{equation}
where $\rho$ is the estimated spatial correlation between the outcome variables
$y_i, y_j, \ldots  \in \vect{y}$, $W$ is a matrix specifying the spatial
relationship between observations $i$ and $j$ in the dataset,
$\vect{\beta}$ is a vector of estimable coefficients relating the attributes $X$
of an observation to its outcome, and $\vect{\gamma}$ is a vector of
estimated coefficients relating the attributes of an observation's {\em neighbors}
to its outcome. If only correlation and not dependence exists in the data
generating process, it may  be more efficient to estimate a restricted
spatial autoregressive model and therefore
save degrees of freedom in the estimation. \citet{tra_autoregression_paper}
illustrate a decision algorithm originally proposed by \citet{Florax2003} which uses
a likelihood ratio test of the SDM against the restricted spatial error model (SEM):
\begin{align}
{\boldsymbol y} &= X{\boldsymbol \beta} + {\boldsymbol u},  {\boldsymbol u} = \lambda W {\boldsymbol u} +
{\boldsymbol \epsilon} \label{eq:SEM}\\
{\boldsymbol y} &= X{\boldsymbol \beta} + (I-\lambda W)^{-1}{\boldsymbol \epsilon}\notag
\end{align}
If the likelihood ratio test rejects that the models are equivalent, then the SDM
should be used because its estimates are unbiased. If the test cannot reject
equivalents, the SEM is more efficient.

@@

@@

We initially estimated potential values of the $\beta$ coefficients by
applying a limited-memory, box-constrained optimization algorithm maximizing the
SDM model log-likehood \cite. The constraint required that the coefficient on
size be positive and that on distance negative; all else equal, we assert people
will prefer to use larger parks and parks that are nearer to their residence. This
resulted in $\beta_d = -1.98$  and $\beta_s = 0.000$. Though the "most likely"
parameter value, in this case it was unsatisfactory for two reasons: first,
our prior assertion that park size is a valuable attribute of a park, and second
the observation that in this specification tracts bordering on the corners of
Central Park were measured with poor overall accessibility. we subsequently adjusted the $\beta_s$ value
so that the

to a value of $1.00$

@@

<<logsum_betas>>=
tracts@data <- tracts@data %>%
  mutate(
    access_ls = calculate_park_logsums(log(distances), parks$log_size,
                                       betas = c(-1.98, 0.3)),
    tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
                                       betas = c(-1.98, 0.3, 0.1)),
    access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
    tweets_ls = (tweets_ls - mean(tweets_ls) )/ sd(tweets_ls)
  )

@@

<<map, echo=FALSE, message=FALSE, warning=FALSE>>=
library(ggmap)
map <- ggmap(get_map(unname(st_bbox(tracts_sf)), source = "stamen", maptype = "toner-lite"),
             extent = "device")
map +
  geom_sf(data = tracts_sf, inherit.aes=FALSE,
          aes(fill = cut(access_ls, breaks = c(-Inf, -3, -2, -1, 0, 2, 3, Inf))),
          alpha = 0.8, color = NA) +
  scale_fill_brewer("Relative Park Logsum", palette = "GnBu")
@
  \caption{Relative surplus-based accessibility in New York City}
  \label{fig:map}
\end{figure*}

\subsection{Results}

<<base_formula>>=
base_formula <- formula(
  ~ log(density) +
    fulltime + college + single +
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
    black + asian + hispanic + other)

<<spatial selection>>=
physact_base_sem <- errorsarlm(update(base_formula,
                                      physact ~ . + highincome + lowincome),
                             data = tracts@data, listw = W, zero.policy = TRUE)
physact_base_sdm <- lagsarlm(update(base_formula,
                                    physact ~ . + highincome + lowincome),
                           data = tracts@data, listw = W, zero.policy = TRUE,
                           type = "mixed")
pa_base_logincome <- update(physact_base_sem, . ~ . +log(income) - highincome - lowincome)
test_sdmsem <- lmtest::lrtest(physact_base_sem, physact_base_sdm)
@

We estimated SDM and SEM models regressing the physical activity rate against
against the base covariates (without the accessibility logsum). A likelihood
ratio test failed to reject that the SEM is equivalent
($\chi^2 = \Sexpr{test_sdmsem$Chisq[2]}$ on \Sexpr{test_sdmsem$Df[2]} degrees of
freedom), so we use the SEM specification only going forward.
Table \ref{tab:pa_models} presents the estimated coefficients for the base
model with income represented in two ways. In the first, the share of the tract
population in
the bottom three income deciles and the share of the population in the top three
income deciles are each included as independent predictors. In the second, the
logarithm of the mean tract income is included as a predictor.
Notwithstanding the higher model likelihood of the deciles model, comparing
coefficients across the models shows that the predictors change magnitude
or significance only marginally, if at all. Overall, the base model
predictors are significant and of intuitive direction. Tracts with higher shares
of full-time workers, college-educated adults, young adults, and high-income
households all show a greater share of individuals engaging in regular physical
activity. Conversely, tracts with greater population density and a greater
share of single adults, children, seniors, minorities of all types, and low
income households all report lower rates of physical activity.

The other two models presented in Table \ref{tab:pa_models} include
the log-sum accessibility statistic described in Section
\ref{subsec:accessibility}  added as a predictor. In both income specifications,
the

base model as
well as models including logsum-based accessibility with and without Twitter
activity. A few initial observations can be made from this table. First, adding
the logsum measurement significantly improves the overall model fit, using a
likelihood ratio test on the model log-likelihood statistics. Second, the
coefficient estimates on the controlling variables remain virtually unchanged
from the base model.

<<access_sdm>>=
# estimate models
pa_access_sem <- update(physact_base_sem, .~ . + access_ls)
pa_access_logincome <- update(pa_access_sem, . ~ . + log(income) - highincome - lowincome)
pa_models <- list("Base" = physact_base_sem,
                  "Base: log(Income)" = pa_base_logincome,
                  "Access" = pa_access_sem,
                  "Access: log(Income)" = pa_access_logincome)

<<screen_pa_models, eval = FALSE>>=
screenreg(pa_models, digits = 4, table = FALSE,
          stars = c(0.001, 0.01, 0.05, 0.1) )
@

\begin{table*}
  \caption{SEM Coefficients Predicting Physical Activity Rates}
  \label{tab:pa_models}
<<pamodelchart, results="asis">>=
termnames <- c("(Intercept)", "log(density)" ,
               "Full-time workers", "College-educated",  "Single Adults",
               "Share of children (0-17)", "Share of young adults (18-29)",
               "Share of seniors (65+)", "Black population share",
               "Asian population share",
               "Hispanic population share", "Other Minorities",
               "Share high income", "Share low income",
               "$\\lambda$", "$\\log($Income$)$",
               "Access logsum")
texreg(pa_models, digits = 4, table = FALSE, single.row = TRUE,
       booktabs = TRUE, use.packages = FALSE,
       custom.coef.names = termnames)
@
\end{table*}

<<obesity_models>>=
obesity_models <- lapply(pa_models, function(model){
  update(update(model, obesity ~ . ))
})

names(obesity_models) <- names(pa_models)
@

<<screen_obese_models, eval = FALSE>>=
screenreg(obesity_models, digits = 4, table = FALSE,
          stars = c(0.001, 0.01, 0.05, 0.1) )
@
We now consider the impacts of a model where the dependent variable is
the {\em obesity} rate, and physical activity becomes an independent covariate
alongside the controlling variables and the accessibility logsum variable.

\begin{table*}
  \caption{SEM Coefficients Predicting Obesity Rates}
  \label{tab:obese_models}
<<obesemodelchart, results="asis">>=
termnames <- c("(Intercept)", "log(density)" ,
               "Full-time workers", "College-educated",  "Single Adults",
               "Share of children (0-17)", "Share of young adults (18-29)",
               "Share of seniors (65+)", "Black population share",
               "Asian population share",
               "Hispanic population share", "Other Minorities",
               "Share high income", "Share low income",
               "$\\lambda$", "$\\log($Income$)$",
               "Access logsum")
texreg(obesity_models, digits = 4, table = FALSE, single.row = TRUE,
       booktabs = TRUE, use.packages = FALSE,
       custom.coef.names = termnames)
@
\end{table*}
